Collocated Deduplication with Fault Isolation for Virtual Machine Snapshot Backup

### 2013-08-27 snapshot_system/arch.tex
revise deletion

In a busy VM cluster, snapshot deletions can occur frequently.
Deduplication complicates the deletion process because space saving relies on the sharing of data
and it requires the global reference of deldted chunks to be identified before  they can be safely removed.
While we can use the mark-and-sweep technique~\cite{Guo2011}, 
it still takes significant resource to conduct this process every time there is a snapshot deletion.
In the case of Alibaba, snapshot backup is conducted automatically and there are 
about 10 snapshot stored for every user. When there is
a new snapshot created every day,  there will be  a snapshot expired everyday to maintain
a balanced storage use. Given the large number of snapshot deletion requests, we seek
a fast solution with a very low resource usage to delete snapshots.
%In a traditional deduplication system, deletion often require looking at global scope to
%resolve the data dependency between logical backup entities and physical data chunks.
Our VM-centric snapshot storage design simplifies the deletion process since 
we can focus on  unreferenced chunks within each VM.
The PDS data chunks are commonly shared all VMs and we do not consider them
during snapshot deletion.  The selection of PDS data chunks is updated periodically independent of snapshot deletion process.
Another resource-saving strategy we propose is
an {\em approximate} deletion strategy to trade deletion accuracy for
speed and resource usage. Our method sacrifices a small percent of storage leakage
to efficiently identify unused chunks in $O(n)$ time, with $n$ being the logical number of non-PDS chunks 
to be deleted from a VM snapshot store.  The algorithm contains three aspects.

%Our system adopts VM-centric snapshot  lazy delete strategy so that all snapshot deletions are scheduled
%in the backup time window at midnight. 
%Therefore, snapshot deletions must be fast enough to fit in time window and
%efficient enough to satisfy our resource constraints.
%However, there is no simple solution can achieve these goals with high reliability.
%Our hybrid deletion strategy, using fuzzy deletion regularly and accurate deletion periodically,
%accomplishs our speed, resource usage and relibility goals very well.
\begin{itemize}
\item {\bf Computation for snapshot fingerprint summary.}
Every time there is a new snapshot created,
we compute a bloom-filter with $z$ bits as the summary of reference pointers for all non-PDS chunks used 
in this snapshot.

@@

\item {\bf Approximate deletion with fast summary comparison.}
When there is a snapshot deletion,  
we need to identify if  chunks to be deleted from that snapshot
are still used by other snapshots. 
This is done approximately and quickly by comparing the 
reference pointers of deleted snapshot with
the merged reference bloom-filter summary of other live snapshots.
The merging of live snapshot bloom-filter bits uses the logical OR operator 
and the merged vector still takes $z$ bits.
Since the number of live snapshots is limited for 
each VM (e.g. 10 in the Alibaba's production system), 
the time cost of this comparison is small.
%Instead of scanning the entire append store indices, we merge the type-1 summaries of all
%valid snapshots. 
%Since each VM has uniform bloom filter parameters to create snapshot summaries, 
%such merged summeries give us a compact representation of
%all block fingerprints that are still in use.
%Thus by the property of bloom filter, if a fingerprint is not found in merged summaries, 
%we are certain that block is no longer used by any valid snapshot, it would be then added
%to append store's deletion log.

If a chunk's reference pointer is not found in the merged summary vector, we are sure that
this chunk is not used by any live snapshots, thus it's safe delete it. 
However, among all the chunks that suppose to be deleted, 
there is a small false-positive ratio which
would identify unused chunks as in use, resulting in temporary storage leakage.

\item {\bf Periodic repair of leakage}
%[exlpain why second bloom filter, why scan append store]
Since there are certain unused chunks which are not deleted during
the approximate deletion, leakage repair is conducted periodically.
In this phase we load the reference pointers of all chunks from a VM snapshot store as a table in memory,
then scan through all the snapshots' segment recipes. For each reference pointer that is in use by a snapshot,
we mark the corresponding entry in the table as in use. Finally those unmarked (and therefore unused) reference pointers in the
table represent chunks that can be safely deleted.

The cost of leak repair mainly come from holding a table of reference pointers and the scan of all snapshots' metadata.
Consider each reference pointer consumes 8 bytes plus 1 byte as mark field, a VM that has 40GB backup data with about
10 million chunks will need 90MB of memory to construct such a table. 
Also, scanning all snapshots metadata is many times slower when compared
to the approximate deletion which only scans single snapshot's metadata.
It's worth mention that compared to previously-developed mark-and-sweep techniques, 
our leak repair still has advantages because the scope of repair is restricted within a single VM's snapshot 
store due to our VM-centric design, and we perform leakage repair periodically.
%We cannot simply repeat the phase 1 multiple times to reduce temporary storage leakage, because:
%\begin{enumerate} 
%\item After several runs of phase-1, it is proven that the merged type-1 summaries cannot sieve remaining unused blocks, due to the false-positive property of bloom filter.
%\item The recipes of deleted snapshots have been removed from the system, thus we are not able to obtain the deleted block fingerprints from any metadata, the only way to discover them is to scan the append store indices.
%\end{enumerate}
\end{itemize}

@@

%{\bf Discussion}
We now estimate the storage leakage and how often leak repair needs to be conducted.
Assuming a VM always maintain $h$ snapshots in the backup, and it creates and deletes one snapshot
everyday. Let $u$ be the total number of chunks brought by the initial backup, $\Delta u$ be the average
number of additional unique chunks added from one snapshot to the next snapshot. Then the total number of unique
chunks used is about:
\[
U = u + (h-1)\Delta u.
\]

Each bloom filter vector has  $z$ bits for each snapshot and let $j$ be the number of hash functions used by the
bloom filter. The probablity that a particular bit is still in all $h$ summary vectors is still 0 is 
$(1- \frac{1}{z}) ^{j U}$. Notice that when a chunk appears multiple times in these summary vectors, it does not 
increase the probability of a particular bit being 0 in all $h$ vectors.
Then the false-positive rate $\epsilon$ is: 
\[
\epsilon = (1-(1-\frac{1}{z})^{jU})^j.
\]

For each snapshot deletion, the amount of chunks need to be deleted is nearly identical to the number of
newly add chunks $\Delta u$. However, some of chunks among them are not detected as unused in our approximation
algorithm, thus forms the storage leakage. Let $R$ be the total number of runs of approximate deletion since
last repair. we estimate  the total leakage $L$ after $R$ runs as:
\[
L = R \Delta u \epsilon
\]

When leakage ratio $L/U$ exceeds a pre-defined threshold $\tau$, we need to execute a leak repair:
\[
\frac{L}{U} = \frac{R \Delta u \epsilon}{u+(h-1)\Delta u } > \tau 
\Longrightarrow R > \frac{\tau}{\epsilon}\times\frac{u + (h-1)\Delta u}{\Delta u}
\]
For example in our tested dataset,  
each VM keeps $h=10$ snapshots and each snapshot has
about 1-5\% of new data. Thus $\frac{\Delta u}{u}=0.05$ at most. With 40GB snapshot size, $u\approx = 10$ millions.
Then $U=10.45$ millions.
We choose  $\epsilon = 0.01$ and $\tau=0.1$. Then from the bloom filter  false positive formula, 
$z=10U=100.45$ million bits. $R=290$.
Then we would expect leak repair be triggered once for 
every 290 runs of approximate deletion. 
When one machine hosts 25 VMs and there is one snapshot deletion per day per VM, there would be 
only one full leak repair for one VM scheduled for every 12 days. Each repairs uses about 90MB memory on average
as discussed earlier and takes a short period of time.
% which is sufficiently small to offset the heavy I/O workload of the mark-sweep process.

>>>>

In a busy VM cluster, snapshot deletions can occur frequently.
Deduplication complicates the deletion process because space saving relies on the sharing of data
and it requires the global reference of deleted chunks to be identified before  they can be safely removed.
While we can use the mark-and-sweep technique~\cite{Guo2011}, 
it still takes significant resource to conduct this process every time there is a snapshot deletion.
In the case of Alibaba, snapshot backup is conducted automatically and there are 
about 10 snapshot stored for every user. When there is
a new snapshot created every day,  there will be  a snapshot expired everyday to maintain
a balanced storage use.

%Given the large number of snapshot deletion requests, 
We seek a fast solution with a low resource usage to delete snapshots.
%In a traditional deduplication system, deletion often require looking at global scope to
%resolve the data dependency between logical backup entities and physical data chunks.
Our VM-centric snapshot storage design simplifies the deletion process since 
we can focus on  unreferenced chunks within each VM.
The PDS data chunks are commonly shared all VMs and we do not consider them
during snapshot deletion.  The selection of PDS data chunks is updated periodically independent of snapshot deletion process.
Another resource-saving strategy we propose is
an {\em approximate} deletion strategy to trade deletion accuracy for
speed and resource usage. Our method sacrifices a small percent of storage leakage
to efficiently identify unused chunks.
% in $O(n)$ time where $n$ is the total number of non-PDS chunks to be deleted from a VM snapshot store.  
The algorithm contains three aspects.

%Our system adopts VM-centric snapshot  lazy delete strategy so that all snapshot deletions are scheduled
%in the backup time window at midnight. 
%Therefore, snapshot deletions must be fast enough to fit in time window and
%efficient enough to satisfy our resource constraints.
%However, there is no simple solution can achieve these goals with high reliability.
%Our hybrid deletion strategy, using fuzzy deletion regularly and accurate deletion periodically,
%accomplishs our speed, resource usage and relibility goals very well.
\begin{itemize}
\item {\bf Computation for snapshot fingerprint summary.}
Every time there is a new snapshot created,
we compute a bloom-filter with $z$ bits as the summary of reference pointers for all non-PDS chunks used 
in this snapshot. Given $h$ snapshots stored for a VM, there are $h$ summary vectors maintained.

@@

\item {\bf Approximate deletion with fast summary comparison.}
When there is a snapshot deletion,  
we need to identify if  chunks to be deleted from that snapshot
are still used by other snapshots. 
This is done approximately and quickly by comparing the 
reference pointers of deleted snapshot with
the merged reference bloom-filter summary of other live snapshots.
The merging of live snapshot bloom-filter bits uses the logical OR operator 
and the merged vector still takes $z$ bits.
Since the number of live snapshots $h$ is limited for
each VM, 
the time and memory cost of this comparison is small, linear to the number of chunks to be deleted.
%Instead of scanning the entire append store indices, we merge the type-1 summaries of all
%valid snapshots. 
%Since each VM has uniform bloom filter parameters to create snapshot summaries, 
%such merged summeries give us a compact representation of
%all block fingerprints that are still in use.
%Thus by the property of bloom filter, if a fingerprint is not found in merged summaries, 
%we are certain that block is no longer used by any valid snapshot, it would be then added
%to append store's deletion log.

If a chunk's reference pointer is not found in the merged summary vector, we are sure that
this chunk is not used by any live snapshots, thus it's safe delete it. 
However, among all the chunks to be deleted, 
there is a small percentage of unused chunks  which
are misjudged as  being in use, resulting in a storage leakage.

\item {\bf Periodic repair of leakage}.
%[exlpain why second bloom filter, why scan append store]
%Since there are certain unused chunks which are not deleted during the approximate deletion, 
Leakage repair is conducted periodically to fix the above approximation error.
This procedure compares the live chunks for each VM with what are truly used through the VM snapshot metadata recipe.
That requires a scanning of all chunks in a VM; however it is a VM-specific procedure and thus
the cost is relatively small compared to the traditional mark-sweep procedure~\cite{?} which scans snapshot 
chunks from all VMs.
For example,
% are live used   we load the reference pointers of all chunks from a VM snapshot store as a table in memory,
%then scan through all the snapshots' segment recipes. For each reference pointer that is in use by a snapshot,
%we mark the corresponding entry in the table as in use. Finally those unmarked (and therefore unused) reference pointers in the
%table represent chunks that can be safely deleted.
%The cost of leak repair mainly come from holding a table of reference pointers and the scan of all snapshots' metadata.
consider each reference pointer consumes 8 bytes plus  1 mark bit, a VM that has 40GB backup data with about
10 million chunks will need less than 90MB of memory to complete a VM-specific mark-sweep process.
%, scanning all snapshots metadata is many times slower when compared
%to the approximate deletion which only scans single snapshot's metadata.
%It's worth mention that compared to previously-developed mark-and-sweep techniques, 
%our leak repair still has advantages because the scope of repair is restricted within a single VM's snapshot 
%store due to our VM-centric design, and we perform leakage repair periodically.
%We cannot simply repeat the phase 1 multiple times to reduce temporary storage leakage, because:
%\begin{enumerate} 
%\item After several runs of phase-1, it is proven that the merged type-1 summaries cannot sieve remaining unused blocks, due to the false-positive property of bloom filter.
%\item The recipes of deleted snapshots have been removed from the system, thus we are not able to obtain the deleted block fingerprints from any metadata, the only way to discover them is to scan the append store indices.
%\end{enumerate}
\end{itemize}

@@

%{\bf Discussion}
We now estimate the size of storage leakage and how often leak repair needs to be conducted,
given  a VM which keeps $h$ snapshots in the backup storage, and it creates and deletes one snapshot
everyday. Let $u$ be the total number of chunks brought by the initial backup, $\Delta u$ be the average
number of additional unique chunks added from one snapshot to the next snapshot version. Then the total number of unique
chunks used in a VM is about:
\[
U = u + (h-1)\Delta u.
\]

Each bloom filter vector has  $z$ bits for each snapshot and let $j$ be the number of hash functions used by the
bloom filter. The probablity that a particular bit is 0  in all $h$ summary vectors is  
$(1- \frac{1}{z}) ^{j U}$. Notice that a chunk may appear multiple times in these summary vectors; however, this should not 
increase the probability of being a 0 bit in all $h$ summary vectors.
Then the misjudgment rate of being in use $\epsilon$ is: 
\begin{equation}
\label{eq:falserate}
\epsilon = (1-(1-\frac{1}{z})^{jU})^j.
\end{equation}

For each snapshot deletion, the amount of chunks need to be deleted is nearly identical to the number of
newly add chunks $\Delta u$. 
%However, some of chunks among them are not detected as unused in our approximation
%algorithm, thus forms the storage leakage. 
Let $R$ be the total number of runs of approximate deletion between two consecutive 
repairs. we estimate  the total leakage $L$ after $R$ runs as:
\[
L = R \Delta u \epsilon
\]

When leakage ratio $L/U$ exceeds a pre-defined threshold $\tau$, we need to execute a leak repair:

\begin{equation}
\label{eq:leakrepair}
\frac{L}{U} = \frac{R \Delta u \epsilon}{u+(h-1)\Delta u } > \tau 
\Longrightarrow R > \frac{\tau}{\epsilon}\times\frac{u + (h-1)\Delta u}{\Delta u}
\end{equation}

For example in our tested dataset,  
each VM keeps $h=10$ snapshots and each snapshot has
about 1-5\% of new data. Thus $\frac{\Delta u}{u} \leq 0.05$. For a 40GB snapshot, $u\approx  10$ millions.
Then $U=10.45$ millions.
We choose  $\epsilon = 0.01$ and $\tau=0.1$.  From Equation~\ref{eq:falserate}, 
$z=10U=100.45$ million bits. From Equation~\ref{eq:leakrepair}, 
leak repair should be triggered once for every R=290 runs of approximate deletion. 
When one machine hosts 25 VMs and there is one snapshot deletion per day per VM, there would be 
only one full leak repair for one VM scheduled for every 12 days. Each repairs uses at most  90MB memory on average
as discussed earlier and takes a short period of time.
% which is sufficiently small to offset the heavy I/O workload of the mark-sweep process.
